---
title: "BikeSharing system"
author: "Xuyang Bai, Mengyan Zhu"
date: "4/30/2018"
output: pdf_document
---
```{r}
library(ggplot2)
library(caret)
require(party)
library(rpart)
library(rpart.plot)
library(MASS)
library(tree)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(ggthemes)
library(ggrepel)
library(gridExtra)
library(lubridate)
library(maptree)
library(glmnet)
library(devtools) 
library(causalTree)
library(zoo)
library(ggfortify)
library(corrplot)
library(plyr)

```
Read data
```{r}
hour <- read.csv("~/Desktop/Bike sharing/hour.csv")
day <- read.csv("~/Desktop/Bike sharing/day.csv")
```

3. Exploratory Analysis
3.1 Overall Understanding
```{r}
summary(day)
```

3.2 Possible Transformation
```{r}
#Distribution
hist(day$casual, breaks = 20, xlab = "casual",main = "casual users")
hist(day$registered, breaks = 20, xlab = "registered",main = "registered users")
hist(day$cnt, breaks = 20, xlab = "count",main = "total users")

hist(log(day$windspeed), breaks = 20, xlab = "windspeed",main = "windspeed")
hist(log(day$atemp), breaks = 20, xlab = "atemp",main = "temperature")
hist(day$hum, breaks = 20, xlab = "hum",main = "humidity")

boxplot(day$casual,main = "casual users")
boxplot(day$windspeed, main = "wind speed")
boxplot(day$hum,main = "humidity")
```

3.3 Missing value and outliers
```{r}
#Missing value
sum(is.na(hour))
sum(is.na(day))

#Outlier
## Working day
ggplot(day, aes(x=factor(workingday) ,y=casual)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
               outlier.size=4)

ggplot(day, aes(x=factor(workingday) ,y=registered)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
               outlier.size=4)

## Season& month& hour
ggplot(hour, aes(x=factor(hr) ,y=cnt)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
               outlier.size=4)
ggplot(day, aes(x=factor(season) ,y=cnt)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
               outlier.size=4)
ggplot(day, aes(x=factor(mnth) ,y=cnt)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
               outlier.size=4)
```

3.4 Find correlations between variables
```{r}
#Correlation between IVs
res=cor(subset(day,select=c(3,4,5,6,7,8,9,10,11,12,13,14,15,16)))
par(mfrow=c(1,1))
corrplot(res, type = "upper", order = "hclust", tl.col = "black",tl.srt = 45)

```

3.4.1 Detail examine of histograms and uncover potential relationships
```{r}
# Season& month& hour
ggplot(hour, aes(x=factor(hr) ,y=cnt)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
               outlier.size=4)
ggplot(day, aes(x=factor(season) ,y=cnt)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
               outlier.size=4)
ggplot(day, aes(x=factor(mnth) ,y=cnt)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
               outlier.size=4)

#Time series
CntDay<-zoo(day$cnt, order.by=as.Date(as.Date(day$dteday), format='%m/%d/%Y'))
CntDay<-ts(CntDay)
class(CntDay)
autoplot(CntDay,ts.colour = 'cyan4') + xlab("Date")+ylab("Total count of bike rentals") + ggtitle("The Time Series of total count")


#Hourly count of rental bike by weekday
day_summary <- ddply(hour,.(weekday,hr),
                     summarise, count = mean(cnt))
ggplot(hour, aes(x = factor(hr), y = count, colour = factor(weekday))) +
  geom_point(data = day_summary, aes(group=weekday)) +
  geom_line(data = day_summary, aes(group=weekday)) +  
  scale_colour_discrete(labels=c("Sun","Mon","Tue","Wed","Thur","Fri","Sat"))+
  guides(color=guide_legend(title=NULL))+
  ggtitle("Hourly count of rental bike by user")+
  scale_x_discrete("Hour") +
  scale_y_continuous("Count") +
  ggtitle("Hourly count of rental bike by weekday")


#Hourly count of rental bike by user
cas<-aggregate(x=hour["casual"], list(time=hour$hr), mean)
reg<-aggregate(x=hour["registered"], list(time=hour$hr), mean)
merge<-merge(cas,reg, all=TRUE)
ggplot(merge, aes(x=hour, y=count)) + 
  geom_point(aes(x=time, y=casual,color="cyan") ) + 
  geom_line(aes(x=time,y=casual, color="cyan",group="casual"))+ 
  geom_point(aes(x=time,y=registered,color="red") ) + 
  geom_line(aes(x=time,y=registered, color="red"))+
  scale_colour_discrete(labels=c("casual","registered"))+
  guides(color=guide_legend(title=NULL))+
  ggtitle("Hourly count of rental bike by user")+theme(legend.position = 'top')

```

3.4.2 The count of rental bikes under different weather condition
```{r}
# Correlation between DV and IV
ggplot(day, aes(atemp,cnt))+geom_point(aes(color=factor(weathersit)), alpha=0.8)+xlab("feeling temperature")+ylab("total rental bikes")+
  ggtitle("Correlation between feeling temperature and total rental bikes") + theme_bw()
ggplot(day, aes(atemp,cnt))+geom_point(aes(color=factor(season)), alpha=0.8)+xlab("feeling temperature")+ylab("total rental bikes")+
  ggtitle("Correlation between feeling temperature and total rental bikes")+ theme_bw()

ggplot(day, aes(hum,cnt))+geom_point(aes(color=factor(weathersit)), alpha=0.8) +xlab("humidity")+ylab("total rental bikes")+
  ggtitle("Correlation between humidity and total rental bikes")+theme_bw()
ggplot(day, aes(hum,cnt))+geom_point(aes(color=factor(season)), alpha=0.8) +xlab("humidity")+ylab("total rental bikes")+
  ggtitle("Correlation between humidity and total rental bikes")+theme_bw()

ggplot(day, aes(windspeed,cnt))+geom_point(aes(color=factor(weathersit)), alpha=0.8) +xlab("windspeed")+ylab("total rental bikes")+
  ggtitle("Correlation between windspeed and total rental bikes")+theme_bw()
ggplot(day, aes(windspeed,cnt))+geom_point(aes(color=factor(season)), alpha=0.8) +xlab("windspeed")+ylab("total rental bikes")+
  ggtitle("Correlation between windspeed and total rental bikes")+theme_bw()

ggplot(day, aes(factor(season),cnt)) + geom_bar(stat="identity",fill="orange")+
  ggtitle("Total rental bikes by season")+xlab("season")+ylab("total rental bikes")
ggplot(day, aes(factor(weekday),cnt)) + geom_bar(stat="identity",fill="orange")+
  ggtitle("Total rental bikes by weekday")+xlab("weekday")+ylab("total rental bikes")

ggplot(day, aes(temp,atemp)) + geom_point(color="cyan4")+
  ggtitle("Temperature and feeling temprature")+xlab("temp")+ylab("atemp")

ggplot(day, aes(factor(weathersit),hum)) + geom_point(color="cyan4")+
  ggtitle("Weathersit and humidity")+xlab("weathersit")+ylab("humidity")

```

3.5 Using Post Double Lasso Selection to resolve omitted variable bias
```{r}
model_z <-model.matrix(~ as.factor(yr)+as.factor(holiday)*as.factor(weekday)*as.factor(workingday)
                        +as.factor(mnth)*as.factor(season)*as.factor(weathersit)*hum*windspeed,
                       data=day)[,-1]


y<-day$cnt
x<-day$atemp

cnt_ols <- lm(y~x+model_z)
summary(cnt_ols)

# LASSO on y
ylasso<-cv.glmnet(model_z,y,alpha=1)
# Coefficients for best (lowest CV MSE) lambda
y.z<-coef(ylasso,ylasso$lambda.min)


# Save coefficients that are non-zero
lasso_y <- as.matrix(y.z)
#get names of each non-zero coeff
y.znames <- rownames(lasso_y)[lasso_y!=0]
y.znames

# LASSO on x (same procedure by on X)
xlasso <- cv.glmnet(model_z,x,alpha=1)
x.z<-coef(xlasso,xlasso$lambda.min)
lasso_x <- as.matrix(x.z)
x.znames<-rownames(lasso_x)[lasso_x!=0]
x.znames
# Combine the variables selected in each LASSO
z <- union(x.znames, y.znames)
z <-z[z!="(Intercept)"]
z
# Combined post-double-LASSO model
model_lasso <-paste("y~x+",paste(z,collapse="+"))

# Reformat model string
mod1 <- strsplit(model_lasso,"")[[1]]
for (i in 1:length(mod1)){
  if (mod1[i]=="("|mod1[i]==")"|mod1[i]==","|mod1[i]=="^"|mod1[i]=="="|(mod1[i]==" "&& i>5)|mod1[i]==":"){
    mod1[i]="."
  }
}
model_lasso2 <-paste(mod1, collapse="")

# Create another data.frame for R to reference
z1 <- cbind(y,x,model_z)
z1 <- data.frame(z1)
# Estimate post-double-LASSO
pd_lasso <-lm(model_lasso2,data=z1)
summary(pd_lasso)
names(coef(pd_lasso))
```

3.6 Using Causal Tree to find Treatment Effect of Feeling Temperature and Working Day	
3.6.1 Select feeling temperature as treatment variable
```{r}
# Estimate causal tree

tree <- causalTree(cnt~ as.factor(yr)+as.factor(holiday)+as.factor(weekday)+as.factor(workingday)
                   +as.factor(mnth)+as.factor(season)+as.factor(weathersit)+hum+windspeed,data=day,treatment = X,
                   split.Rule = "CT", cv.option = "CT", split.Honest = T, cv.Honest = T, split.Bucket = F, 
                   xval = 5, cp = 0, minsize = 15, propensity = 0.5)
opcp <- tree$cptable[,1][which.min(tree$cptable[,4])]
opfit <- prune(tree, opcp)

# Plot and summarize causal tree
rpart.plot(opfit)
summary(opfit)
draw.tree(opfit,cex=3)
```

3.6.2 Select working day as treatment variable
```{r}
X1<-day$workingday

tree1 <- causalTree(cnt~ as.factor(yr)+as.factor(holiday)+as.factor(weekday)+atemp
                   +as.factor(mnth)+as.factor(season)+as.factor(weathersit)+hum+windspeed,data=day,treatment =X1,
                   split.Rule = "CT", cv.option = "CT", split.Honest = T, cv.Honest = T, split.Bucket = F, 
                   xval = 5, cp = 0, minsize = 20, propensity = 0.5)
opcp1 <- tree1$cptable[,1][which.min(tree$cptable[,4])]
opfit1 <- prune(tree1, opcp1)
rpart.plot(opfit1)
summary(opfit1)

median(day$atemp)
day$AtempC<-ifelse(day$atemp>0.486733,1,0)
day$AtempC
```

4. Modeling
4.1 Predicting total number of bike rental(cnt) mainly in SPSS 
Split train test
```{r}
set.seed(3456)
trainIndex <- createDataPartition(hour$cnt, p = .7, 
                                  list = FALSE, 
                                  times = 1)
hourTrain <- hour[ trainIndex,]
hourTest  <- hour[-trainIndex,]
names(hour)
```
Predict total count use Simple regression tree 
```{r}
cart <- tree(cnt~factor(season)+factor(yr)+factor(mnth)+factor(hr)+factor(holiday)+factor(weekday)+factor(workingday)+factor(weathersit)+temp+atemp+hum+windspeed , data = hourTrain)
summary(cart)
plot(cart)
text(cart,pretty=0)
cv.P=cv.tree(cart)
summary(cv.P)
prune.P=prune.tree(cart,best=5)
plot(prune.P)
text(prune.P,pretty=0)
yhat=predict(cart,newdata=hourTest)
print(MSETree<-mean((yhat-day$cnt)^2))
```

4.2 Clusterting mainly in SPSS
```{r}
# 3 Cluster for hour 
library(readxl)
hour_cluster <- read_excel("hour,cluster.xlsx")
#FM<-na.omit(FM)
d <- dist(hour_cluster, method = "euclidean")
fit <- hclust(d, method="ward.D")
plot(fit)
rect.hclust(fit, k=3, border="red")
```

4.3 Time series analysis
Use time series to predict future 1 day total count/registered/casual
```{r}
library(readxl)
library(zoo)
library(ggfortify)
CntDay <- read_excel("~/Desktop/Bike sharing/total count per day.xlsx")
CntDay  <-zoo(CntDay$cnt, order.by=as.Date(as.Date(CntDay$dteday), format='%m/%d/%Y'))
CntDay<-ts(CntDay)
class(CntDay)
autoplot(CntDay,ts.colour = 'green') + xlab("Date")+ylab("Total count of bike rentals") + ggtitle("The Time Series of total count")
```
periodic, 24 hour seasonal
ACF: AR exist for seasonal and nonseasonal
PACF: seasonal is only related last 1 day
SARIMA200 24
SARIMA100 24

Total count
```{r}
library(readxl)
library(astsa)
hourTS <- read_excel("hourTS.xlsx")
View(hourTS)
Cnt <- ts(hourTS$cnt,frequency = 24)
acf2(Cnt)
sarima
SARIMA200<-sarima(Cnt, 2,0,0,1,0,0,24)
SARIMA300<-sarima(Cnt, 3,0,0,1,0,0,24)
#Best model
SARIMA100<-sarima(Cnt, 1,0,1,4,0,1,24)
summary(SARIMA100)
Forecast<-sarima.for(Cnt,24,1,0,1,4,0,1,24)
Forecast
```

```{r}
library(Hmisc)
lm1<-lm(Cnt~Lag(Cnt,1)+Lag(Cnt,24)+hourTS$atemp+as.factor(hourTS$yr)+as.factor(hourTS$holiday)*as.factor(hourTS$weekday)*as.factor(hourTS$workingday)+as.factor(hourTS$mnth)*as.factor(hourTS$season)*as.factor(hourTS$weathersit)*hourTS$hum*hourTS$windspeed,data=hourTS)
summary(lm1)
```
casual
```{r}
casual <- ts(hourTS$casual,frequency = 24)
acf2(casual)
#Best model
SARIMA100<-sarima(casual, 1,0,2,3,0,2,24)
```
```{r}
summary(SARIMA100)
Forecast<-sarima.for(casual,24,1,0,2,3,0,2,24)
Forecast
```
```{r}
lm2<-lm(casual~Lag(Cnt,1)+Lag(Cnt,24)+hourTS$atemp+as.factor(hourTS$yr)+as.factor(hourTS$holiday)*as.factor(hourTS$weekday)*as.factor(hourTS$workingday)+as.factor(hourTS$mnth)*as.factor(hourTS$season)*as.factor(hourTS$weathersit)*hourTS$hum*hourTS$windspeed,data=hourTS)
summary(lm2)
```
registered
```{r}
registered <- ts(hourTS$registered,frequency = 24)
acf2(registered)
```



